---
title: "Women in STEM at UT"
output: html_notebook
---

Using UT's Diversity tracker, we are able to display the relations of women in STEM fields at UT.

We using `packrat` to keep track of the packages that were used to do this simple analysis.

```{r}
# install.packages('tidyverse') install tidyverse to do reading + ploting of data, etc.
library(tidyverse)
```

```{r}
enrollment.2016.fall.mf <- read_csv('data/stem_enrollment_male_female_fall_2016.csv')
glimpse(enrollment.2016.fall.mf)
```

We need to convert data to tidyset so that we can do some plotting with it.

```{r}
enrollment.2016.fall.mf.tidy <- enrollment.2016.fall.mf %>%
  select(-Total) %>%
  gather(key = "Gender", value = "Population", Male:Female)
```


```{r}
stem_colors = c("Female" = "#754A7e", "Male" = "#2FBEBD", "Nonbinary" = "#58595B")

enrollment.2016.fall.mf.stem.plot <- enrollment.2016.fall.mf.tidy %>%
  select(Major,Population,Gender) %>%
  group_by(Major) %>%
    ggplot(data = ., aes(x = Major, y = Population, group = Gender, fill = Gender)) +
  scale_fill_manual(values = stem_colors) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    width = 1,
    alpha = 1
  ) +
  ylab('Percentage') +
  xlab('Major') +
  scale_y_continuous(breaks = seq(0,1000, by = 200), limits = c(0,1000), expand = c(0, 0))+#scale_y_continuous(breaks = seq(0,600, by = 200), limits = c(0,600), expand = c(0, 0))+
  #scale_x_continuous(breaks = seq(0,12, by = 3), limits = c(0,12.5), expand = c(0, 0))+
  coord_flip()+
  guides(col = guide_legend(nrow = 2)) +
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(color="black", size = 1, fill = NA),
    axis.ticks.length = unit(-0.3, "lines"),
    axis.text.y = element_text(color="black", size = 6*text_scaling, margin = margin(l = -7, r = 6, unit = "pt")),
    axis.text.x = element_text(color="black", size = 10*text_scaling, margin = margin(t = 6, b = -7, unit = "pt")),
    axis.title.y = element_text(color="black", size = 10*text_scaling, margin = margin(l = -1, r = 8, unit = "pt"), angle = 90),
    axis.title.x = element_text(color="black", size = 10*text_scaling, margin = margin(t = 8, b = -1, unit = "pt")),
    axis.ticks.y = element_line(color="black", size = 1),
    axis.ticks.x = element_line(color="black", size = 1),
    plot.margin = margin(t = 10, r = 12, b = 7, l = 7, unit = "pt"),
    panel.grid.major.x = element_line(color="#eeeeee", size = 1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(color="black", size = 10*text_scaling),
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(color="black", size = 10*text_scaling),
    legend.key.height = unit(12, "pt"),
    legend.key.width = unit(16, "pt"),
    legend.margin = margin(0, unit = "pt")
  )
enrollment.2016.fall.mf.stem.plot 
```

```{r}
ggsave(plot = enrollment.2016.fall.mf.stem.plot, filename = 'plots/enrollment.2016.fall.mf.stem.plot.svg', width = 4.15, height = 3, units = 'in')
```





```{r}
library(stringr)
enrollment.2010_2016.tidy <- enrollment.2010_2016 %>%
  gather(key = "Population.Type", value = "Population", 4:22) %>%
  separate(Population.Type, into=c("Semester", "Year", "Gender"))

enrollment.2010_2016.tidyr <- enrollment.2010_2016.tidy %>%
  filter(!(Gender == "All")) %>%
  group_by(School, Major, Level, Year) %>%
  mutate(Population.Pct = Population / sum(Population))

```

We'll start plotting the ratios of STEM fields at UT

## Gender Ratio in STEM Majors at UT
```{r}
stem_colors = c("Female" = "#754A7e", "Male" = "#2FBEBD", "Nonbinary" = "#58595B")
text_scaling = 1.25

enrollment.2010_2016.plot <- enrollment.2010_2016.tidyr %>%
    ggplot(data = ., aes(x = Major, y = Population, group = Major, color = Gender)) +
  scale_color_manual(values = stem_colors) +
  geom_jitter(
    aes(size = )
  )+
  geom_line(
    size = 2,
    alpha = 1
  ) +
  ylab('# of students') +
  xlab('Major') +
  scale_y_continuous(breaks = seq(0,600, by = 150), limits = c(0,600), expand = c(0, 0))+#scale_y_continuous(breaks = seq(0,600, by = 200), limits = c(0,600), expand = c(0, 0))+
  #scale_x_continuous(breaks = seq(0,12, by = 3), limits = c(0,12.5), expand = c(0, 0))+
  #scale_size(guide="none")+
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(color="black", size = 1, fill = NA),
    axis.ticks.length = unit(-0.3, "lines"),
    axis.text.y = element_text(color="black", size = 9*text_scaling, margin = margin(l = -7, r = 6, unit = "pt")),
    axis.text.x = element_text(color="black", size = 9*text_scaling, margin = margin(t = 6, b = -7, unit = "pt")),
    axis.title.y = element_text(color="black", size = 10*text_scaling, margin = margin(l = -1, r = 8, unit = "pt"), angle = 90),
    axis.title.x = element_text(color="black", size = 10*text_scaling, margin = margin(t = 8, b = -1, unit = "pt")),
    axis.ticks.y = element_line(color="black", size = 1),
    axis.ticks.x = element_line(color="black", size = 1),
    plot.margin = margin(t = 10, r = 12, b = 7, l = 7, unit = "pt"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(color="black", size = 10*text_scaling),
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(color="black", size = 10*text_scaling),
    legend.key.height = unit(12, "pt"),
    legend.key.width = unit(16, "pt"),
    legend.margin = margin(0, unit = "pt")
  )
enrollment.2010_2016.plot
#ggsave(plot = enrollment.2010_2016.plot, filename = 'plots/enrollment.2010_2016.plot.svg', width = 4.15, height = 3, units = 'in')
```

# Investigate numbers by origin

```{r}
enrollment.2016.fall <- read_csv('data/stem_enrollment_female_fall_2016.csv')

enrollment.2016.fall.tidy <- enrollment.2016.fall %>%
  gather(key = "Race", value = "Population", Hispanic:White) %>%
  mutate(Population.pct = Population / Total) %>%
  select(-Total)
```

```{r}
enrollment.2016.fall.tidy %>%
  select(Race, Population) %>%
  group_by(Race) %>%
  summarise(Population = sum(Population)) %>%
  mutate(Population.pct = Population / sum(Population))
```



```{r}
# Overall Race plot
enrollment.2016.fall.plot <- enrollment.2016.fall.tidy %>%
  select(Race, Population) %>%
  group_by(Race) %>%
  summarise(Population = sum(Population)) %>%
  mutate(Population.pct = Population / sum(Population)) %>%
    ggplot(data = ., aes(x = Race, y = Population, group = Race, fill = Race)) +
  scale_fill_brewer(palette = "Set2") +
  geom_bar(
    stat = "identity",
    size = 2,
    alpha = 1
  ) +
  ylab('female undergraduates in STEM at Tennessee') +
  xlab('Race') +
  scale_y_continuous(breaks = c(200,3600), limits = c(0,4000), expand = c(0, 0))+#scale_y_continuous(breaks = seq(0,600, by = 200), limits = c(0,600), expand = c(0, 0))+
  #scale_x_continuous(breaks = seq(0,12, by = 3), limits = c(0,12.5), expand = c(0, 0))+
  #scale_size(guide="none")+
  coord_flip()+
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(color="black", size = 1, fill = NA),
    axis.ticks.length = unit(-0.3, "lines"),
    axis.text.y = element_text(color="black", size = 9*text_scaling, margin = margin(l = -7, r = 6, unit = "pt")),
    axis.text.x = element_text(color="black", size = 9*text_scaling, margin = margin(t = 6, b = -7, unit = "pt")),
    axis.title.y = element_text(color="black", size = 10*text_scaling, margin = margin(l = -1, r = 8, unit = "pt"), angle = 90),
    axis.title.x = element_text(color="black", size = 10*text_scaling, margin = margin(t = 8, b = -1, unit = "pt")),
    axis.ticks.y = element_line(color="black", size = 1),
    axis.ticks.x = element_line(color="black", size = 1),
    plot.margin = margin(t = 10, r = 12, b = 7, l = 7, unit = "pt"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(color="black", size = 10*text_scaling),
    strip.background = element_blank(),
    legend.position = "none"#,
    # legend.text = element_text(color="black", size = 10*text_scaling),
    # legend.key.height = unit(12, "pt"),
    # legend.key.width = unit(16, "pt"),
    # legend.margin = margin(0, unit = "pt")
  )
enrollment.2016.fall.plot 
```

```{r}
ggsave(plot = enrollment.2016.fall.plot, filename = 'plots/enrollment.2016.fall.plot.svg', width = 4.15, height = 3, units = 'in')
```

```{r}
# WoC in STEM fields
enrollment.2016.fall.woc.stem.plot <- enrollment.2016.fall.tidy %>%
  select(Major,Race, Population.pct) %>%
  filter(!(Race == "White")) %>%
  group_by(Major) %>%
    ggplot(data = ., aes(x = Major, y = Population.pct*100, group = Race, color = Race)) +
  scale_color_brewer(palette = "Set2") +
  geom_jitter(
    size = 3,
    alpha = 1
  ) +
  ylab('Percentage') +
  xlab('Major') +
  scale_y_continuous(breaks = seq(0,20, by = 5), limits = c(0,20), expand = c(0, 0))+#scale_y_continuous(breaks = seq(0,600, by = 200), limits = c(0,600), expand = c(0, 0))+
  #scale_x_continuous(breaks = seq(0,12, by = 3), limits = c(0,12.5), expand = c(0, 0))+
  #scale_size(guide="none")+
  guides(col = guide_legend(nrow = 2)) +
  theme(
    panel.background = element_blank(),
    panel.border = element_rect(color="black", size = 1, fill = NA),
    axis.ticks.length = unit(-0.3, "lines"),
    axis.text.y = element_text(color="black", size = 9*text_scaling, margin = margin(l = -7, r = 6, unit = "pt")),
    axis.text.x = element_text(color="black", size = 6*text_scaling, angle = 90, margin = margin(t = 6, b = -7, unit = "pt")),
    axis.title.y = element_text(color="black", size = 10*text_scaling, margin = margin(l = -1, r = 8, unit = "pt"), angle = 90),
    axis.title.x = element_text(color="black", size = 10*text_scaling, margin = margin(t = 8, b = -1, unit = "pt")),
    axis.ticks.y = element_line(color="black", size = 1),
    axis.ticks.x = element_line(color="black", size = 1),
    plot.margin = margin(t = 10, r = 12, b = 7, l = 7, unit = "pt"),
    panel.grid.major.x = element_line(color="#eeeeee", size = 1),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(color="black", size = 10*text_scaling),
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(color="black", size = 10*text_scaling),
    legend.key.height = unit(12, "pt"),
    legend.key.width = unit(16, "pt"),
    legend.margin = margin(0, unit = "pt")
  )
enrollment.2016.fall.woc.stem.plot 
```

```{r}
ggsave(plot = enrollment.2016.fall.woc.stem.plot, filename = 'plots/enrollment.2016.fall.woc.stem.plot.svg', width = 4.15, height = 5, units = 'in')
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
