---
title: "Women in STEM at UT"
output: html_notebook
---

Using UT's Diversity tracker, we are able to display the relations of women in STEM fields at UT.

We using `packrat` to keep track of the packages that were used to do this simple analysis.

```{r}
# install.packages('tidyverse') install tidyverse to do reading + ploting of data, etc.
library(tidyverse)
```

```{r}
enrollment.2016.fall <- read_csv('data/enrollment_fall_2016.csv')
glimpse(enrollment.2016.fall)
```

We need to convert data to tidyset so that we can do some plotting with it.

```{r}
enrollment.2016.fall.tidy <- enrollment.2016.fall %>%
  gather(Population.Type, Population, All.Total:Nonresident_alien.Women)
```

```{r}
#install.packages("stringr")
library(stringr)
enrollment.2016.fall.tidyr <- enrollment.2016.fall.tidy %>%
  mutate(Race = str_match(Population.Type, "(\\w+)\\.")[,2],
         Gender = str_match(Population.Type, "\\.(\\w+)")[,2]) %>%
  select(-Population.Type)
```


We'll start plotting the ratios of STEM fields at UT

## Gender Ratio in STEM Fields at UT
```{r}
enrollment.2016.fall <- enrollment.2016.fall %>%
  mutate(Percentage = sum(Female))
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
